---
title: "Prevalence and reporting practices of a priori power analyses"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2025-01-21"
---

# Load packages

```{r packages, include = FALSE, echo=FALSE}

library(here)
library(dplyr)
library(tidyr)
library(kableExtra)
```

# Load data

```{r data}
source(here("r_scripts", "data_analysis", "02_clean_data.R"))
```

# Results

## Prevalence of a priori power analyses

```{r prevalence, echo=FALSE}
# Two studies included both an a priori and post-hoc power analysis. 
# This can be checked at raw_data.xlsx (see comments section)
(prevalence_power <- cleaned_data %>% 
  group_by(power_analysis) %>%
  summarise(n = n()) %>%
  mutate(n = ifelse(power_analysis == "post-hoc", n + 2, n)) %>%
  mutate(prop = round(n / sum(n) *100, 0)))
```

# Reporting practices of a priori power analyses

## Estimated sample size

```{r sample_size, echo = FALSE}
(power_sample <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   summarise(n = sum(!is.na(power_sample)),
            prop = round(n / prevalence_power$n[[1]]*100, 0)))
```

## Dependent variable

```{r dv, echo=FALSE}
(dv <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(specified_dv_power) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0)))
```

## ES magnitude

```{r magnitude_es, echo=FALSE}
(magnitude_es <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   summarise(total = n(),
             reported = sum(!is.na(magnitude_es)),
             prop = round(reported / total *100,0)))
```

## ES type

```{r es_type, echo=FALSE}
(type_es <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(type_es) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0)))
```

## ES type and magnitude

```{r es, echo=FALSE}
(es <- cleaned_data %>% 
   filter(power_analysis == "a priori" & 
            !is.na(magnitude_es) & 
            type_es == "yes, standardised") %>%
   group_by(type_standardised_es) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0)))
```

## Type Cohen's d ES

```{r type_cohen echo=FALSE}
cleaned_data %>% 
   filter(power_analysis == "a priori" & 
            !is.na(magnitude_es) & 
            type_es == "yes, standardised" &
            type_standardised_es == "Cohen's d family") %>%
   group_by(type_cohen) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0))
```

## Type Hedges'g ES

```{r type_hedge, echo=FALSE}
cleaned_data %>% 
   filter(power_analysis == "a priori" & 
            !is.na(magnitude_es) & 
            type_es == "yes, standardised" &
            type_standardised_es == "Hedge's g family") %>%
   group_by(type_hedge) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0))
```

## ES justification

```{r es_justification, echo=FALSE}
(es_justification <- cleaned_data %>% 
   filter(power_analysis == "a priori" & !is.na(magnitude_es)) %>%
   select(es_justification) %>%
   mutate(es_justification = 
           case_when(es_justification %in% c("meta-analysis and previous studies",
                                             "meta-analysis and previous study",
                                             "previous studies",
                                             "two meta-analyses") ~ "unclear",
                     es_justification %in% c("methodological paper", 
                                              "position statement", 
                                              "systematic review") ~ "others",
                     is.na(es_justification) ~ "not reported",
                     TRUE ~ es_justification)) %>%
  group_by(es_justification) %>%
  summarise(n = n()) %>%
  mutate(prop = round(n / sum(n)*100, 0)) %>%
  arrange(desc(n)) %>%
  mutate(merged = paste0(prop, "%", " (", n, ")")) %>% 
  select(es_justification, merged) %>%
  kable(col.names = c("ES Justification", "Prevalence"), 
        align = c("l", "c")))
```

## Statistical test

```{r test, echo=FALSE}

# by test
(test <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(statistical_test_power) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n/sum(n)*100, 0))) %>%
   arrange(desc(n))

# by family of tests 
(family_test <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   mutate(test = case_when(
     statistical_test_power %in% c("paired t-test", "unpaired t-test") ~ "t-test",
     statistical_test_power == "not reported" ~ "not reported",
     statistical_test_power == "unclear" ~ "unclear",
     TRUE ~ 'F-test')) %>%
  group_by(test) %>%
  summarise(n = n()) %>%
  mutate(prop = round(n/sum(n)*100, 0)))

```

## Intended statistical power

```{r intended_power, echo = FALSE}

(intended_power <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
  mutate(intended_power = case_when(intended_power == 0.8 ~ "0.8",
                                    intended_power == "not reported" ~ "not reported",
                                    intended_power %in% c(0.81, 0.85, 0.9, 0.95, 0.99) ~ "Other")) %>%
  group_by(intended_power) %>% 
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n/sum(n)*100, 0)))
```

## Software

```{r software, echo=FALSE}

(software <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(software) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n/sum(n)*100, 0)))
```

## Alpha level

```{r alpha, echo=FALSE}
(alpha <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>% 
   mutate(alpha_level = case_when(alpha_level %in% c("assumed two-tailed at 0.05",
                                                    "two-tailed at 0.05") ~ "0.05",
                                 alpha_level %in% c("assumed two-tailed other",
                                                    "two-tailed other") ~ "other",
                                 TRUE ~ "not reported")) %>%
   group_by(alpha_level) %>% 
   summarise(n = n()) %>%
   mutate(prop = round(n / sum(n)*100, 0)))
```

## Sidedness of t-test

```{r sidedness, echo=FALSE}
(sidedness <- cleaned_data %>% 
   filter(power_analysis == "a priori" & 
          statistical_test_power %in% c("paired t-test", "unpaired t-test")) %>%
  group_by(alpha_level) %>% 
  summarise(n = n()) %>%
  mutate(prop = round(n / sum(n)*100, 0)))
```

## Table 1

| Parameter | Prevalence |
|-------------------------------------------|-----------------------------|
| Primary outcome |  |
| Specific primary outcome | `r paste0(dv[3,3], "%", " (", dv[3,2], ")")` |
| Umbrella term | `r paste0(dv[4,3], "%", " (", dv[4,2], ")")` |
| Unclear | `r paste0(dv[2,3], "%", " (", dv[2,2], ")")` |
| Not reported | `r paste0(dv[1,3], "%", " (", dv[1,2], ")")` |
| ES type |  |
| Unstandardised | `r paste0(type_es[3,3], "%", " (", type_es[3,2], ")")` |
| Standardised | `r paste0(type_es[2,3], "%", " (", type_es[2,2], ")")` |
| Not reported | `r paste0(type_es[1,3], "%", " (", type_es[1,2], ")")` |
| ES magnitude | `r paste0(magnitude_es[1,3], "%", " (", magnitude_es[1,2], ")")` |
| Statistical test |  |
| t-test | `r paste0(family_test[3,3], "%", " (", family_test[3,2], ")")` |
| ANOVA | `r paste0(family_test[1,3], "%", " (", family_test[1,2], ")")` |
| Unclear | `r paste0(family_test[4,3], "%", " (", family_test[4,2], ")")` |
| Not reported | `r paste0(family_test[2,3], "%", " (", family_test[2,2], ")")` |
| Sidedness of the test |  |
| Two-sided | `r paste0(sidedness[2,3], "%", " (", sidedness[2,2], ")")` |
| Not reported | `r paste0(sidedness[1,3], "%", " (", sidedness[1,2], ")")` |
| Alpha level |  |
| 0.05 | `r paste0(alpha[1,3], "%", " (", alpha[1,2], ")")` |
| Other | `r paste0(alpha[3,3], "%", " (", alpha[3,2], ")")` |
| Not reported | `r paste0(alpha[2,3], "%", " (", alpha[2,2], ")")` |
| Intended Power |  |
| 80% | `r paste0(intended_power[1,3], "%", " (", intended_power[1,2], ")")` |
| Other | `r paste0(intended_power[2,3], "%", " (", intended_power[2,2], ")")` |
| Not reported | `r paste0(intended_power[3,3], "%", " (", intended_power[3,2], ")")` |
| Software |  |
| G\*Power | `r paste0(software[1,3], "%", " (", software[1,2], ")")` |
| Other | `r paste0(software[2,3], "%", " (", software[2,2], ")")` |
| Not reported | `r paste0(software[3,3], "%", " (", software[3,2], ")")` |

## A priori power analyses reporting nº of measurements and correlation when a one-way within-subject ANOVA

```{r echo=FALSE}

cleaned_data %>% 
  filter(statistical_test_power == "one-way within-subject ANOVA") %>%
  summarise(n = n(),
            reported = sum(!is.na(measurements) & !is.na(correlation)),
            prop = reported / n*100)

```

## A priori power analyses reporting nº of measurements, groups and correlation when a two-way mixed ANOVA

```{r echo=FALSE}

cleaned_data %>% 
  filter(statistical_test_power %in% c("two-way mixed ANOVA", 
                                       "two-way within-subject ANOVA",
                                       "two-way mixed ANOVA",
                                       "two-way within-subject MANOVA")) %>%
  summarise(n = n(),
            reported = sum(!is.na(measurements)  & 
                           !is.na(groups) & 
                           !is.na(correlation)),
            prop = reported / n*100)

```

## A priori power analyses reporting nº groups and conditions when a two-way between-subject ANOVA

```{r echo=FALSE}

cleaned_data %>% 
  filter(statistical_test_power == "one-way between-subject ANOVA") %>%
  summarise(n = n(),
            reported = sum(!is.na(measurements) &
                    !is.na(groups)),
            prop = reported / n*100)

```
