---
title: "Computational reproducibility of a priori power analyses"
date: "2025-01-21"
---

```{r echo=FALSE, warning = FALSE, message = FALSE}

# Load packages
library(here)
library(readxl)
library(dplyr)

# Load data
source(here("r_scripts", "data_analysis", "02_clean_data.R"))

```

# Proportion of a priori power analyses that could be reproduced

```{r reproducibility}

(reproducibility <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(reproducibility) %>%
   summarise(n = n()) %>%
   mutate(prop = ceiling(n / base::sum(n)*100)))

```

# Proportion of unstandardised effect sizes that could be reproduced

```{r unstand_es}

(unstand_es <- cleaned_data %>% 
  filter(power_analysis == "a priori") %>%
  filter(type_es == "yes, unstandardised") %>%
  group_by(reproducibility) %>%
  summarise(n = n()) %>%
  mutate(prop = n / base::sum(n)*100))

```

# Number of guessed parameters when a priori power analysis where "conditionally reproducible"

```{r guesses}

(guesses <- cleaned_data %>%
   filter(power_analysis == "a priori" &
          reproducibility == "conditionally reproducible") %>% 
   group_by(guessed_parameters) %>%
   summarise(n = n()) %>%
   mutate(prop = round(n / base::sum(n) * 100,0)))

```

Estimates of computationally reproducibility are illustrated in Figure 1. `r reproducibility[2,3]`% (`r reproducibility[2,2]`/`r sum(data$power_analysis == "a priori")`) of a priori power analyses were deemed fully non-reproducible because they did not include an effect size to begin with. Thus, we attempted to reproduce a total of `r sum(data$power_analysis == "a priori")-reproducibility[2,2]` a priori power analysis. `r reproducibility[3,3]`% (`r reproducibility[3,2]`/`r sum(data$power_analysis == "a priori")`) were deemed non-reproducible and `r reproducibility[4,3]` (`r reproducibility[4,2]`/`r sum(data$power_analysis == "a priori")`) were deemed reproducible. `r reproducibility[1,3]`% (`r reproducibility[1,2]`/`r sum(data$power_analysis == "a priori")`) were deemed conditionally reproducible after making assumptions for the missing parameters required to conduct an a priori power analysis. Of these, `r guesses[2,3]`% (`r guesses[2,3]`/`r reproducibility[1,2]`) required to make more than one assumption including the statistical test, number of groups and/or measurements and the correlation between measurements.

# Proportion of a priori power analyses that used G\*Power's default option

```{r default_gpower, echo=FALSE}

(default <-  cleaned_data %>%
   filter(power_analysis == "a priori") %>%
   mutate(es = case_when(type_standardised_es %in% c("eta partial squared (ηp^2)",
                                                     "eta squared (η^2)") ~ "eta",
                                                     TRUE ~ type_standardised_es)) %>%
   filter(es %in% c("Cohen's f", "eta")) %>%
   filter(es_justification %in% c("previous study",
                                  "previous studies",
                                  "meta_analysis") & 
          statistical_test_gpower %in% c("within-subject ANOVA",
                                         "mixed ANOVA") &
          reproducibility %in% c("reproducible",
                                 "conditionally reproducible") &
          software == "G*Power") %>%
   summarise(total = length(gpower_default_option), 
             yes_default = base::sum(gpower_default_option == "yes", na.rm = TRUE),
             prop_yes_default = yes_default / total * 100))

```

```{r echo=FALSE}
# Here we do the same as above but we do not consider the ES
cleaned_data %>%
  filter(power_analysis == "a priori") %>%
  filter(statistical_test_gpower %in% c("within-subject ANOVA","mixed ANOVA") &
         reproducibility %in% c("reproducible", "conditionally reproducible") &    
         software == "G*Power") %>%
  summarise(total = length(gpower_default_option), 
            yes_default = base::sum(gpower_default_option == "yes", na.rm = TRUE),
            prop_yes_default = yes_default / total * 100)
```

# Proportion of a priori power analyses using G\*Power

```{r g_power}
(gpower <- cleaned_data %>%
  filter(power_analysis == "a priori" & software == "G*Power") %>%
  summarise(n = n()) %>%
  mutate(prop = round((n / sum(cleaned_data$power_analysis == "a priori"))*100, 0)))
```

# Proportion of a priori power analyses using within-subject or mixed ANOVA

```{r prop_ws_anova}
cleaned_data %>%
  filter(statistical_test_gpower %in% c("within-subject ANOVA","mixed ANOVA")) %>%
  summarise(n = n()) %>%
  mutate(prop = round((n / sum(cleaned_data$power_analysis == "a priori"))*100, 0))

cleaned_data %>%
  filter(power_analysis == "a priori") %>%
  mutate(statistical_test_power = case_when(
    statistical_test_power %in% c("one-way within-subject ANOVA",
                                  "two-way mixed ANOVA",
                                  "two-way within-subject ANOVA",
                                  "two-way within-subject MANOVA") ~ 
      "within-subject ANOVA", 
    TRUE ~ statistical_test_power)) %>%
  group_by(statistical_test_power) %>%
  summarise(n = n())
```
