---
title: "Computational reproducibility of a priori power analyses"
output:
  word_document: default
  html_document: default
date: "2025-01-21"
---

```{r, include=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(here)
library(readxl)
library(dplyr)
library(tidyr)

# Load data
source(here("r_scripts", "data_analysis", "02_clean_data.R"))

```

# Proportion of a priori power analyses that could be reproduced
``` {r echo=FALSE}

(reproducibility <- cleaned_data %>% 
   filter(power_analysis == "a priori") %>%
   group_by(reproducibility) %>%
   summarise(n = n()) %>%
   mutate(prop = ceiling(n / base::sum(n)*100)))

```

# Proportion of unstandardised effect sizes that could be reproduced
``` {r echo=FALSE}

cleaned_data %>% 
  filter(power_analysis == "a priori") %>%
  filter(type_es == "yes, unstandardised") %>%
  group_by(reproducibility) %>%
  summarise(n = n()) %>%
  mutate(prop = n / base::sum(n)*100)

```

# Number of guessed parameters when a priori power analysis where "conditionally reproducible"
```{r echo=FALSE}

(guesses <- cleaned_data %>%
   filter(power_analysis == "a priori" &
          reproducibility == "conditionally reproducible") %>% 
   group_by(guessed_parameters) %>%
   summarise(n = n()) %>%
   mutate(prop = round(n / base::sum(n) * 100,0)))

```
Estimates of computationally reproducibility are illustrated in Figure 1. 
`r reproducibility[2,3]`% (`r reproducibility[2,2]`/`r sum(data$power_analysis == "a priori")`) 
of a priori power analyses were deemed fully non-reproducible because they did not 
include an effect size to begin with. Thus, we attempted to reproduce a total of 
`r sum(data$power_analysis == "a priori")-reproducibility[2,2]` a priori power 
analysis. `r reproducibility[3,3]`% (`r reproducibility[3,2]`/`r sum(data$power_analysis == "a priori")`) 
were deemed non-reproducible and `r reproducibility[4,3]` (`r reproducibility[4,2]`/`r sum(data$power_analysis == "a priori")`) were deemed reproducible. `r reproducibility[1,3]`% (`r reproducibility[1,2]`/`r sum(data$power_analysis == "a priori")`) were deemed conditionally reproducible after making assumptions 
for the missing parameters required to conduct an a priori power analysis. Of these, 
`r guesses[2,3]`% (`r guesses[2,3]`/`r reproducibility[1,2]`) required to 
make more than one assumption including the statistical test, number of groups 
and/or measurements and the correlation between measurements. 


# Proportion of a priori power analyses that used G*Power's default option
```{r echo=FALSE}

(default <-  cleaned_data %>%
   filter(power_analysis == "a priori") %>%
   mutate(es = case_when(type_standardised_es %in% c("eta partial squared (ηp^2)",
                                                     "eta squared (η^2)") ~ "eta",
                                                     TRUE ~ type_standardised_es)) %>%
   filter(es %in% c("Cohen's f", "eta")) %>%
   filter(es_justification %in% c("previous study",
                                  "previous studies",
                                  "meta_analysis") & 
          statistical_test_gpower %in% c("within-subject ANOVA",
                                         "mixed ANOVA") &
          reproducibility %in% c("reproducible",
                                 "conditionally reproducible") &
          software == "G*Power") %>%
   summarise(total = length(gpower_default_option), 
             yes_default = base::sum(gpower_default_option == "yes", na.rm = TRUE),
             prop_yes_default = yes_default / total * 100))

```
`r default[1]` a priori power analyses that were deemed as "conditionally reproduced" 
or "reproduced" reported a within-subject or mixed ANOVA, used a Cohen’s f or 
η_p^2 obtained from previous study and used GPower. Of those, `r default[3]`%
(`r default[2]`/`r default[1]`) selected the default option (i.e., “as in G*Power 3.0”).


```{r echo=FALSE}

# Here we do the same as above but we do not consider the ES
cleaned_data %>%
  filter(power_analysis == "a priori") %>%
  filter(statistical_test_gpower %in% c("within-subject ANOVA","mixed ANOVA") &
         reproducibility %in% c("reproducible", "conditionally reproducible") &    
         software == "G*Power") %>%
  summarise(total = length(gpower_default_option), 
            yes_default = base::sum(gpower_default_option == "yes", na.rm = TRUE),
            prop_yes_default = yes_default / total * 100)
  
```

# Proportion of a priori power analyses reporting a within-subject or mixed ANOVA
```{r}
cleaned_data %>%
  filter(statistical_test_power %in% c("within-subject ANOVA","mixed ANOVA")) %>%
  length()

```

# Proportion of studies reporting an a priori power analysis and a 
# within-subject or mixed ANOVA
```{r}
cleaned_data %>%
  filter(power_analysis == "a priori" &
         statistical_test_study %in% c("within-subject ANOVA","mixed ANOVA")) %>%
  length()

```

