---
title: "Differences in sample sizes and proportion of studies supporting the hypothesis"
output: html_document
date: "2025-01-21"
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load packages
library(here)
library(readxl)
library(dplyr)
library(ggplot2)

# Load function descriptive_stats()
source(here("r_scripts", "data_analysis", "03_descriptive_stats_function.R")) 

# Load data
source(here("r_scripts", "data_analysis", "02_clean_data.R"))

```


```{r, echo=FALSE}

# Descriptive stats of all sample sizes
data2 %>% 
  select(study_sample) %>%
  pull() %>% 
  descriptive_stats() 

## Descriptive stats of sample sizes without outliers
data2 %>% 
  filter(study_sample < 150) %>% 
  select(study_sample) %>%
  pull() %>% 
  descriptive_stats() 

```


# Differences in sample sizes between studies with and without a priori power analysis
# (with outliers)
```{r echo=FALSE}

# Leave out studies where type of power analysis was unclear
data_poisson <- data2 %>%
  filter(!pre_p == "unclear")

# Checking assumption of variance = mean
data_poisson %>% 
  summarise(average = mean(study_sample),
            variance = var(study_sample))

# The variance is larger than the mean what indicates that there is overdispersion
# Because under a Poisson model we would expect the means and variances of the 
# response to be about the same in various groups, we need to adjust for overdispersion. 

# Change level of reference to no a priori power analysis
data_poisson$pre_p <- relevel(data_poisson$pre_p, ref = 2) 

# Poisson regression
poisson_results <- glm(study_sample ~ pre_p, 
                       family = quasipoisson, 
                       data = data_poisson)

model_summary <- summary(poisson_results)
exp_coefficients <- exp(coef(poisson_results))
exp_ci <- exp(confint(poisson_results))
  
sample_descriptive <- data_poisson %>% 
    group_by(pre_p) %>%
    summarise(average = mean(study_sample),
    sd = sd(study_sample))

```
Studies that reported an a priori power analysis had a sample size of 
`r sample_descriptive[2,2]` ± `r sample_descriptive[2,3]` and those which did not 
had a sample size of `r sample_descrpitive[1,2]` ± `r sample_descriptive[1,3]`. 
The result of the Poisson simple regression indicates that the inclusion of 
an a priori power analysis did not significantly increase sample sizes 
(mean difference = `r exp_coefficients[2]`; 95% CI [`r exp_ci[2,1]``
; `r exp_ci[2,2]`]; p = `r model_summary$coefficients[2,4]`).


# Differences in sample sizes between studies with and without a priori power analysis
# (without outliers)
```{r echo=FALSE}
# Leave out studies where type of power analysis was unclear
data_poisson_2 <- data2 %>% 
  filter(!pre_p == "unclear" &
         study_sample < 150)

# Checking assumption of variance = mean
data_poisson_2 %>% 
  summarise(average = mean(study_sample),
            variance = var(study_sample))

# The variance is larger than the mean what indicates that there is overdispersion
# Because under a Poisson model we would expect the means and variances of the 
# response to be about the same in various groups, we need to adjust for overdispersion. 

# Change level of reference to no a priori power analysis
data_poisson_2$pre_p <- relevel(data_poisson_2$pre_p, ref = 2) 

# Poisson regression
poisson_results_2 <- glm(study_sample ~ pre_p, 
                         family = quasipoisson, 
                         data = data_poisson_2) 

summary_model_2 <- summary(poisson_results_2) # summary statistics of the glm model
exp_coef_2 <- exp(coef(poisson_results_2)) # Exponentiate coefficient for interpretation 
exp_ci_2 <- exp(confint(poisson_results_2)) # Exponentiate 95% CI for interpretation

# descriptive stats
sample_descrip_2 <- data_poisson_2 %>% 
  group_by(pre_p) %>%
  summarise(average = mean(study_sample),
  sd = sd(study_sample))

```
Studies that reported an a priori power analysis had a sample size of 
`r sample_descrip_2[2,2]` ± `r sample_descrip_2[2,3]` and those which did not 
had a sample size of `r sample_descrip_2[1,2]` ± `r sample_descrip_2[1,3]`. 
The result of the Poisson simple regression indicates that the inclusion of an a priori 
power analysis did not significantly increase sample sizes (mean difference = `r exp_coef_2[2]`; 
95% CI [`r exp_ci_2[2,1])`; `r exp_ci_2[2,2])`]; p = `r summary_model_2$coefficients[2,4]`).


# Differences in the proportion of hypothesis supported between studies with and 
# without a priori power analysis
```{r echo=FALSE, warning = FALSE}

# Proportion of studies supporting the hypothesis with and without an a priori 
# power analysis

proportion <- data2 %>% 
  filter(type_hypothesis == "effect") %>%
  group_by(pre_p) %>%
  summarise(support_h = sum(support_hypothesis == "yes"),
            total = n())

proportion_results <- prop.test(x = proportion$support_h,  
                                n = proportion$total)

format(proportion_results$statistic^2, scientific = FALSE)

```
`r proportion_results$estimate[1]`% (`r proportion[1,2]`/`proportion[1,3]`) 
of the studies that used an a priori power analysis reported a significant result 
that supported the hypothesis tested, whereas `r proportion_results$estimate[2]`% (`r proportion[2,2]`/`r proportion[2,3]`)) of the studies 
that did not use an a priori power analysis reported a significant result that 
supported the hypothesis tested. The result of the two-sided two-proportion z test 
indicates that the proportion of studies finding support in favour of the hypothesis 
tested was not statistically different between the studies that used an a priori
power analysis and those studies which did not 
(X-squared = `proportion_results$statistic`, 95% CI of the proportion difference [`proportion_results$conf.int[1]`; `r proportion_results$conf.int[1]`], p = `r proportion_results$p.value`). 
